---
title: "Assignment Lab I"
author: "Melvin Disla"
output:
  pdf_document: default
  word_document: default
---
**Download this Source**

```{r}
source("http://www.openintro.org/stat/data/arbuthnot.R")
arbuthnot
dim(arbuthnot)
names(arbuthnot)
arbuthnot$boys
```

**Exercise 1 What command would you use to extract just the counts of girls baptized? Try it!**

```{r}
arbuthnot$girls
plot(x = arbuthnot$year, y = arbuthnot$girls)
plot(x = arbuthnot$year, y = arbuthnot$girls, type = "l")
?plot
```

**Exercise 2 Is there an apparent trend in the number of girls baptized over the years? How would you describe it?**

```{r}
plot(x = arbuthnot$year, y = arbuthnot$girls, main="Baptism records for girls born in London 1629 to 1710",
xlab="year", ylab = "number of newborn girls",type="l")
5218 + 4683
arbuthnot$boys + arbuthnot$girls
plot(arbuthnot$year, arbuthnot$boys + arbuthnot$girls, type = "l")
5218 / 4683
arbuthnot$boys / arbuthnot$girls
5218 / (5218 + 4683)
arbuthnot$boys / (arbuthnot$boys + arbuthnot$girls)
```

**Exercise 3 Now; make a plot of the proportion of boys over time. What do you see? Tip: If you use the up and down arrow keys, you can scroll through your previous commands, your so-called command history. You can also access it by clicking on the history tab in the upper right panel. This will save you a lot of typing in the future.**

```{r}
boygirlrat<-arbuthnot$boys/(arbuthnot$boys + arbuthnot$girls)
plot(arbuthnot$year, boygirlrat,type="l")
(arbuthnot$boys > arbuthnot$girls)
```

## On Your Own

**In the previous few pages, you recreated some of the displays and preliminary analysis of Arbuthnot’s baptism data. Your assignment involves repeating these steps, but for present day birth records in the United States. Load up the present day data with the following command.**

```{r}
source("http://www.openintro.org/stat/data/present.R")
```


**1.What years are included in this data set? What are the dimensions of the data frame and what are the variable or column names?**

```{r}
present$year
dim(present)
names(present)
```

**2. How do these counts compare to Arbuthnot’s? Are they on a similar scale?**

```{r}
mean(arbuthnot$boys + arbuthnot$girls)
mean(present$boys + present$girls)
```

*We can see that the difference is very big. The counts are on a different scale.*

**3. Make a plot that displays the boy-to-girl ratio for every year in the data set. What do you see? Does Arbuthnot’s observation about boys being born in greater proportion than girls hold up in the U.S.? Include the plot in your response.**

Calculate proportion
```{r}
boysProportion <- present$boys / (present$boys + present$girls)
```

Make plot
```{r}
plot(present$year, boysProportion, type = "l", ylim=c(0.5, 0.53))
```

*Yes it does hold up in the USA.*

**4. In what year did we see the most total number of births in the U.S.? You can refer to the help files or the [R reference card] (http://cran.r-project.org/doc/contrib/Short-refcard.pdf) to find helpful commands.**

```{r}
present$year[(present$boys + present$girls) == max(present$boys + present$girls)]
```


