---
title: "Assignment Lab 2"
author: "Melvin Disla"
date: "5/28/2017"
output: word_document
---

## Getting Started
```{r}
source("http://www.openintro.org/stat/data/cdc.R")
names(cdc)
```

** Excercise 1: How many cases are there in this data set? How many variables? For each variable, identify its data type (e.g. categorical, discrete).**
```{r}
head(cdc)
```

*and similarly we can look at the last few by typing*
##Summaries and tables
```{r}
summary(cdc$weight)
```
*R also functions like a very fancy calculator. If you wanted to compute the interquartile range for the respondents’ weight, you would look at the output from the summary command above and then enter*
```{r}
190 - 140
```
*R also has built-in functions to compute summary statistics one by one. For instance, to calculate the mean, median, and variance of weight, type*
```{r}
mean(cdc$weight) 
var(cdc$weight)
median(cdc$weight)
```
*While it makes sense to describe a quantitative variable like weight in terms of these statistics, what about categorical data? We would instead consider the sample frequency or relative frequency distribution. The function table does this for you by counting the number of times each kind of response was given. For example, to see the number of people who have smoked 100 cigarettes in their lifetime, type*
```{r}
table(cdc$smoke100)
table(cdc$smoke100)/20000
barplot(table(cdc$smoke100))
```
*Notice what we’ve done here! We’ve computed the table of cdc$smoke100 and then immediately applied the graphical function, barplot. This is an important idea: R commands can be nested. You could also break this into two steps by typing the following:*
```{r}
smoke <- table(cdc$smoke100)
barplot(smoke)
```
**Exercise 2 Create a numerical summary for height and age, and compute the interquartile range for each. Compute the relative frequency distribution for gender and exerany. How many males are in the sample? What proportion of the sample reports being in excellent health?**
```{r}
table(cdc$gender,cdc$smoke100)
mosaicplot(table(cdc$gender,cdc$smoke100))
```
**Exercise 3 What does the mosaic plot reveal about smoking habits and gender?**
##Interlude: How R thinks about data
```{r}
dim(cdc)
cdc[567,6]
names(cdc)
1:10
cdc[1:10,]
cdc$weight[567]
cdc$weight[1:10]
```
##A little more on subsetting
```{r}
mdata <- subset(cdc, cdc$gender == "m")
head(mdata)
m_and_over30 <- subset(cdc, gender == "m" & age > 30)
m_or_over30 <- subset(cdc, gender == "m" | age > 30)
```
**Exercise 4 Create a new object called under23_and_smoke that contains all observations of respondents under the age of 23 that have smoked 100 cigarettes in their lifetime. Write the command you used to create the new object as the answer to this exercise.**
##Quantitative data

```{r}
boxplot(cdc$height)
summary(cdc$height)
boxplot(cdc$height ~ cdc$gender)
```
*BMI is a weight to height ratio and can be calculated as:*
```{r}
bmi <- (cdc$weight / cdc$height^2) * 703
boxplot(bmi ~ cdc$genhlth)
```
**Exercuse 5 What does this box plot show? Pick another categorical variable from the data set and see how it relates to BMI. List the variable you chose, why you might think it would have a relationship to BMI, and indicate what the figure seems to suggest.**

```{r}
hist(cdc$age)
hist(bmi)
hist(bmi, breaks = 50)
```
##On Your Own

**1. Make a scatterplot of weight versus desired weight. Describe the relationship between these two variables.**
```{r}
plot(cdc$weight, cdc$wtdesire, xlab = "Weight", ylab = "Desired Weight", las = 1)
```
*There appears to be a general positive correlation between the two variables (as one increases, so does the other), although there is a great deal of spread.*
**2. Let’s consider a new variable: the difference between desired weight (wtdesire) and current weight (weight). Create this new variable by subtracting the two columns in the data frame and assigning them to a new object called wdiff.**
```{r}
cdc$wdiff <- cdc$wtdesire - cdc$weight
```
**3. What type of data is wdiff? If an observation wdiff is 0, what does this mean about the person’s weight and desired weight. What if wdiff is positive or negative?**
*wdiff is a scale variable, but it is discrete since the weights are integer values, and thus it can only take integer values itself. When wdiff is equal to 0 that implies that the person is at their desired weight. If wdiff > 0 then their weight is below their desired value, while wdiff < 0 implies that their actual weight is above their desired value.*
**4. Describe the distribution of wdiff in terms of its center, shape, and spread, including any plots you use. What does this tell us about how people feel about their current weight?**
```{r}
summary(cdc$wdiff)
hist(cdc$wdiff, breaks = 50)
```
*The histogram shows that the distribution is heavily peaked in the region of wdiff between 0 and -20, and the summary values also indicate a median wdiff value of -10. This shows that the majority of people are above their desired weight.*
**5. Using numerical summaries and a side-by-side box plot, determine if men tend to view their weight differently than women.**
```{r}
summary(cdc$wdiff[cdc$gender == "m"])
summary(cdc$wdiff[cdc$gender == "f"])
boxplot(cdc$wdiff ~ cdc$gender, outline = F, ylab = "wdiff")
```
**6. Now it’s time to get creative. Find the mean and standard deviation of weight and determine what proportion of the weights are within one standard deviation of the mean.**
```{r}
mean_weight <- mean(cdc$weight)
mean_weight
sd_weight <- sd(cdc$weight) 
sd_weight
weights_within_one_sd <- subset(cdc, weight < mean_weight + sd_weight & weight > mean_weight - + sd_weight)
dim(weights_within_one_sd)
```
* This is a proportion of 14152/2000 = 0.07076 or 71%. So 71% of peoples weights are in 1 standard deviation of the mean.*
















