---
title: 'Assignment: Lab V'
author: "Melvin Disla"
date: "6/18/2017"
output: word_document
---

## Inference for numerical data

## North Carolina births

*In 2004, the state of North Carolina released a large data set containing information on births recorded in this state. This data set is useful to researchers studying the relation between habits and practices of expectant mothers and the birth of their children. We will work with a random sample of observations from this data set.*

## Exploratory analysis

*Load the nc data set into our workspace.*

```{r}
download.file("http://www.openintro.org/stat/data/nc.RData", destfile = "nc.RData")
load("nc.RData")
```

*We have observations on 13 different variables, some categorical and some numerical. The meaning of each variable is as follows.*

variable	description
fage	father’s age in years.
mage	mother’s age in years.
mature	maturity status of mother.
weeks	length of pregnancy in weeks.
premie	whether the birth was classified as premature (premie) or full-term.
visits	number of hospital visits during pregnancy.
marital	whether mother is married or not married at birth.
gained	weight gained by mother during pregnancy in pounds.
weight	weight of the baby at birth in pounds.
lowbirthweight	whether baby was classified as low birthweight (low) or not (not low).
gender	gender of the baby, female or male.
habit	status of the mother as a nonsmoker or a smoker.
whitemom	whether mom is white or not white.

**Exercise 1:What are the cases in this data set? How many cases are there in our sample?** The cases are the relations between the births of children in North Carolina and the habits of their expectant mothers. There are 1,000 cases.

*As a first step in the analysis, we should consider summaries of the data. This can be done using the summary command:*

```{r}
summary(nc)
```

*As you review the variable summaries, consider which variables are categorical and which are numerical. For numerical variables, are there outliers? If you aren’t sure or want to take a closer look at the data, make a graph.*

*Consider the possible relationship between a mother’s smoking habit and the weight of her baby. Plotting the data is a useful first step because it helps us quickly visualize trends, identify strong associations, and develop research questions.*

**Excercise 2: Make a side-by-side boxplot of habit and weight. What does the plot highlight about the relationship between these two variables?*

```{r}
plot(nc$habit, nc$weight, horizontal=T)
with(nc, plot(weight ~ habit))
smoker = subset(nc, habit=="smoker")
nonsmoker = subset(nc, habit=="nonsmoker")
t.test(smoker$weight, nonsmoker$weight)
```

The plot highlights that overall, nonsmoking mothers gave birth to babies that weighed more. Nonsmoking mothers also had a larger range in baby birth weights than non-smoking mothers did.

*The box plots show how the medians of the two distributions compare, but we can also compare the means of the distributions using the following function to split the weight variable into the habit groups, then take the mean of each using the mean function.*

```{r}
by(nc$weight, nc$habit, mean)
```

*There is an observed difference, but is this difference statistically significant? In order to answer this question we will conduct a hypothesis test .*

## Inference

**Exercise 3: Check if the conditions necessary for inference are satisfied. Note that you will need to obtain sample sizes to check the conditions. You can compute the group size using the same by command above but replacing mean with length.**

```{r}
by(nc$weight, nc$habit, length)
hist(smoker$weight)
hist(nonsmoker$weight)
```

# left skew in both but fairly normal. also let me check via quantile plots:

```{r}
qqnorm(smoker$weight)
qqline(smoker$weight)
qqnorm(nonsmoker$weight)
qqline(nonsmoker$weight)
```

**Exercise 4: Write the hypotheses for testing if the average weights of babies born to smoking and non-smoking mothers are different.**
Ho: There is no difference in the average weights of babies born to smoking mothers and babies born to non-smoking mothers.
Ha: On average, babies born to smoking mothers weigh less than babies born to non-smoking mothers.
*Next, we introduce a new function, inference, that we will use for conducting hypothesis tests and constructing confidence intervals.*
```{r}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```

*Let’s pause for a moment to go through the arguments of this custom function. The first argument is y, which is the response variable that we are interested in: nc$weight. The second argument is the explanatory variable, x, which is the variable that splits the data into two groups, smokers and non-smokers: nc$habit. The third argument, est, is the parameter we’re interested in: "mean" (other options are "median", or "proportion".) Next we decide on the type of inference we want: a hypothesis test ("ht") or a confidence interval ("ci"). When performing a hypothesis test, we also need to supply the null value, which in this case is 0, since the null hypothesis sets the two population means equal to each other. The alternative hypothesis can be "less", "greater", or "twosided". Lastly, the method of inference can be "theoretical" or "simulation" based.*

**Exercise 5: Change the type argument to "ci" to construct and record a confidence interval for the difference between the weights of babies born to smoking and non-smoking mothers.**
By default the function reports an interval for (μnonsmoker−μsmoker) . We can easily change this order by using the order argument:
```{r}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical", 
          order = c("smoker","nonsmoker"))
```

## On your own
**1.** *Calculate a 95% confidence interval for the average length of pregnancies (weeks) and interpret it in context. Note that since you’re doing inference on a single population parameter, there is no explanatory variable, so you can omit the x variable from the function.*
```{r}
meanWeeks <- mean(nc$weeks, na.rm=TRUE)
meanWeeks
sdWeeks <- sd(nc$weeks, na.rm=TRUE)
sdWeeks
n <- nrow(nc$weeks)
n
seWeeks <- sqrt(sdWeeks / n)
seWeeks
inference(y=nc$weeks, est="mean", type="ci", method="theoretical")
```

I am 95% confident that the population mean weeks is between 38.1528 and 38.5165.

**2.** *Calculate a new confidence interval for the same parameter at the 90% confidence level. You can change the confidence level by adding a new argument to the function: conflevel = 0.90.*
```{r}
inference(y=nc$weeks, est="mean", type="ci", method="theoretical", conflevel=0.90)
```

I am 90% confident that the population mean weeks is between 38.182 and 38.4873.

**3.** *Conduct a hypothesis test evaluating whether the average weight gained by younger mothers is different than the average weight gained by mature mothers.*
```{r}
inference(y=nc$weight, x=nc$mature, type="ht", est="mean", 
          null=0, method="theoretical", alternative="twosided")
```

**Due to the p-value being so high, at 0.8526, I fail to reject the null hypothesis. I conclude that the weight of babies born to younger moms is not different than born to mature moms.**

**4.** *Now, a non-inference task: Determine the age cutoff for younger and mature mothers. Use a method of your choice, and explain how your method works.*
```{r}
max(nc[nc$mature == "younger moms",]$mage)
min(nc[nc$mature == "mature moms",]$mage)
```

In this case, younger mom max is 34 and mature mom age is 35, so it seems that the cutoff age for younger mom's is less than or equal to 34, and mature mom's is greater than 34.

**5.** *Pick a pair of numerical and categorical variables and come up with a research question evaluating the relationship between these variables. Formulate the question in a way that it can be answered using a hypothesis test and/or a confidence interval. Answer your question using the inference function, report the statistical results, and also provide an explanation in plain language.*
Does the number of average visits different between younger and mature moms?

Ho: The average of visits equals the same between younger and mature moms.
Ha: The average of visits does not equal the same between younger and mature moms.

```{r}
inference(y=nc$visits, x=nc$mature, type="ht", est="mean", 
          null=0, method="theoretical", alternative="twosided")
```
Given the p-value =  0.15, I fail to reject the null hypothesis and conclude that the average number of visits to the doctor is different between younger and mature moms.
